### 2024-09-23 final_paper/paper.tex
Initial Commit

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  letterpaper,
  DIV=11,
  numbers=noendperiod]{scrartcl}

\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else  
    % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{5}
% Make \paragraph and \subparagraph free-standing
\makeatletter
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}{
    \@ifstar
      \xxxParagraphStar
      \xxxParagraphNoStar
  }
  \newcommand{\xxxParagraphStar}[1]{\oldparagraph*{#1}\mbox{}}
  \newcommand{\xxxParagraphNoStar}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}{
    \@ifstar
      \xxxSubParagraphStar
      \xxxSubParagraphNoStar
  }
  \newcommand{\xxxSubParagraphStar}[1]{\oldsubparagraph*{#1}\mbox{}}
  \newcommand{\xxxSubParagraphNoStar}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
\makeatother

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother

\KOMAoption{captions}{tableheading}
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother

\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage{bookmark}

\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={STA302 Fall 2024 Final Project Part 1: Research Proposal and Data Introduction},
  pdfauthor={Wendy Huang; Gadiel David Flores},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\title{STA302 Fall 2024 Final Project Part 1: Research Proposal and Data
Introduction}
\usepackage{etoolbox}
\makeatletter
\providecommand{\subtitle}[1]{% add subtitle to \maketitle
  \apptocmd{\@title}{\par {\large #1 \par}}{}{}
}
\makeatother
\subtitle{Research Proposal and Data Introduction}
\author{Wendy Huang \and Gadiel David Flores}
\date{2024-04-10}

\begin{document}
\maketitle

\section{Potential Research
Questions}\label{potential-research-questions}

\subsection{Business: How do price, customer reviews, and service
features influence customer satisfaction and repeat business in the
online retail
industry?}\label{business-how-do-price-customer-reviews-and-service-features-influence-customer-satisfaction-and-repeat-business-in-the-online-retail-industry}

\subsection{Studying: How do different study techniques and time
management practices affect academic performance among university
students?}\label{studying-how-do-different-study-techniques-and-time-management-practices-affect-academic-performance-among-university-students}

\subsection{Electric Cars: Are electric vehicles more environmentally
sustainable than gasoline-powered cars when accounting for the energy
sources used for electricity
generation?}\label{electric-cars-are-electric-vehicles-more-environmentally-sustainable-than-gasoline-powered-cars-when-accounting-for-the-energy-sources-used-for-electricity-generation}

\section{Work Division}\label{work-division}

\subsection{Group Member 1:}\label{group-member-1}

\textbf{Main Focus: Research and Data Analysis}

\begin{itemize}
\tightlist
\item
  \textbf{Contributions}:

  \begin{itemize}
  \tightlist
  \item
    Locate and clean the dataset.
  \item
    Write the \textbf{Data Description} section (300 words).
  \item
    Fit the \textbf{Multiple Linear Regression Model} and write the
    \textbf{Preliminary Results} section (300 words).
  \item
    Conduct \textbf{Residual Analysis} and create the required residual
    plots.
  \end{itemize}
\end{itemize}

\subsubsection{Specific Tasks:}\label{specific-tasks}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  \textbf{Locate the dataset} that fits the project requirements
  (open-source, contains a response variable and at least 9 predictors).
\item
  \textbf{Clean the data}: Handle missing values, prepare numerical and
  categorical variables, and ensure it's ready for regression.
\item
  \textbf{Write the Data Description}:

  \begin{itemize}
  \tightlist
  \item
    State where the data was found.
  \item
    Describe how the data was originally collected by the curator.
  \item
    Summarize the response variable and justify why it's suitable for
    linear regression.
  \item
    Summarize the predictor variables, provide numerical/graphical
    summaries, and interpret the descriptive statistics.
  \end{itemize}
\item
  \textbf{Fit the multiple linear regression model} using R,
  incorporating at least 5 predictors, including one categorical
  predictor.
\item
  \textbf{Conduct residual analysis}:

  \begin{itemize}
  \tightlist
  \item
    Check for violations of linear regression assumptions.
  \item
    Create residual plots and interpret their results.
  \end{itemize}
\item
  \textbf{Write the Preliminary Results}:

  \begin{itemize}
  \tightlist
  \item
    Report on the fitted model.
  \item
    Discuss model assumptions (whether violated or not).
  \item
    Compare the preliminary model results to findings from the
    literature (summarized by Group Member 2).
  \end{itemize}
\end{enumerate}

\subsection{Group Member 2:}\label{group-member-2}

\textbf{Main Focus: Research Question, Literature Review, and Report
Writing}

\begin{itemize}
\tightlist
\item
  \textbf{Contributions}:

  \begin{itemize}
  \tightlist
  \item
    Write the \textbf{Introduction} section (350 words).
  \item
    Conduct the \textbf{Literature Review}: Summarize three
    peer-reviewed academic papers related to the research question.
  \item
    Write the \textbf{Ethics Discussion} section (100-200 words).
  \item
    Compile the \textbf{Bibliography} and format citations.
  \item
    Final editing and formatting of the proposal.
  \end{itemize}
\end{itemize}

\subsubsection{Specific Tasks:}\label{specific-tasks-1}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  \textbf{Formulate the research question}: Define a clear, focused
  research question based on the chosen dataset and explain why it's
  suitable for linear regression.
\item
  \textbf{Conduct the literature review}:

  \begin{itemize}
  \tightlist
  \item
    Find and summarize three peer-reviewed academic papers related to
    the topic.
  \item
    Describe how each paper relates to the proposed research question
    and supports the use of linear regression.
  \end{itemize}
\item
  \textbf{Write the Introduction}:

  \begin{itemize}
  \tightlist
  \item
    Introduce the relevance/importance of the research topic.
  \item
    Clearly state the research question.
  \item
    Summarize the results of the three academic papers and explain how
    they connect to your research.
  \item
    Justify why linear regression is an appropriate tool for this
    analysis.
  \end{itemize}
\item
  \textbf{Write the Ethics Discussion}:

  \begin{itemize}
  \tightlist
  \item
    Assess whether the dataset is trustworthy and ethically collected
  \end{itemize}
\item
  \textbf{Compile the bibliography}:

  \begin{itemize}
  \tightlist
  \item
    Ensure proper formatting and inclusion of references in the proposal
    using APA format.
  \end{itemize}
\item
  \textbf{Final editing and formatting}:

  \begin{itemize}
  \tightlist
  \item
    Review the entire document
  \end{itemize}
\end{enumerate}

\section{Introduction (350 words)}\label{introduction-350-words}

Provide an introduction to your research topic, clearly state your
research question, and summarize three peer-reviewed research papers.
Justify why linear regression is appropriate for your research question.

\section{Data Description (300 words)}\label{data-description-300-words}

\subsection{Data Source}\label{data-source}

\subsection{Response and Predictor
Variables}\label{response-and-predictor-variables}

\section{Ethics Discussion (100-200
words)}\label{ethics-discussion-100-200-words}

\section{Preliminary Results (300
words)}\label{preliminary-results-300-words}

\subsection{Model Fitting}\label{model-fitting}

\subsection{Residual Analysis}\label{residual-analysis}

\section{Bibliography}\label{bibliography}

\end{document}