### 2024-12-06 final_paper/paper.qmd

---
title: "STA302 Fall 2024 Final Project Part 1: Research Proposal and Data Introduction"
author: 
- "Wendy Huang"
- "Gadiel David Flores"

date: "07/10/24"
subtitle: "Research Proposal and Data Introduction"
output: pdf_document
format: pdf
number-sections: true
bibliography: references.bib
---
# Introduction
Housing affordability and price fluctuations have become pressing issues in metropolitan areas, and Toronto is no exception.Over the past decade, housing prices have seen a steep upward trend, driven by a combination of housing supply and demand, economic, and policy-related factors.(@5causesbehindtoronto) This study aims to investigate the relationship between several key housing market indicators—namely, the number of available homes, homes sold, homes starting construction, new homes completed, and interest rate hikes—and housing prices in Toronto. By analyzing these factors, this research will contribute to a more comprehensive understanding of Toronto's housing market dynamics.

The rapid rise in housing prices has been widely debated, with supply constraints being one of the leading factors. For example, Glaeser, E. L., Gyourko, J., & Saks, R. E. (@whyhavehousingprices) employed linear regression to analyze the relationship between supply constraints and prices, showing an imbalance between supply and demand results in price hikes.

Additional studies support the notion that housing availability and construction trends influence prices. In their paper, Goodman, A. C., & Thibodeau, T. G. (@metropolitanhousing) explored the linear relationship between new housing construction and price stability in urban areas. The authors found that increased housing completions have a stabilizing effect on prices, especially in high-demand areas.

The paper by Guiwen Bai and Yu Ma (@fundamentalhousing) uses a multi-factor regression model to study how these fundamentals, particularly the cost of borrowing and income tax impacts and interest rates, relate to housing price movements in Canada. The paper confirmed that low interest rates contribute to increased affordability, encouraging more people to enter the housing market.

Given these findings, linear regression is an appropriate tool for analyzing how these supply-side continuous variables interact to affect prices in Toronto. The clear quantitative relationship between supply predictors (available homes, homes sold, etc.) and price response aligns dynamically well with the assumptions of linear regression, making it a suitable method for this analysis. Linear regression allows us to quantify the relationship between these factors and price, providing insights into which variables have the most significant impact.

# Data Description

## Data Source

Using data from Statistics Canada, we extracted five datasets. @StatCan-NewHousingPriceIndex served as the foundation for this research paper, as it contained the main statistic we wanted to analyze. @StatCan-NewHousingPriceIndex represents the new housing price index, which calculates the average cost of new housing each month by multiplying by 100 and dividing by the base year. This provides an index reflecting housing costs for any given year and month. We then utilized @StatCan-CPI, @StatCan-GDP, @StatCan-HousingStarts, @StatCan-FinancialMarketStats, and @StatCan-HousingAbsorptions to extract our 13 predictors, including Absorption, GDP, CPI, interest rates, and Construction. Specifically, Absorption and Construction are further divided into two parts. Absorbtion is divided into semi-detached homes and single homes as well as empty and sold houses for any given month. Construction is categorized into three parts: new construction, under construction, and finished construction. These datasets were cleaned by removing missing values, converting data into whole integers, and merging them into one dataset. Additionally, the data spans from 1997 to 2016 and is divided quarterly, providing 76 entries. Lastly, we used interest rates to create our categorical predictor. This was done by comparing the entry i with entry i+1 in order to determine if there was an interest hike.

## Response and Predictor Variables
Given our research question how do supply-side factors, such as housing availability, sales volume, construction starts, completions, and interest rate, impact housing prices in Toronto, the response variable chosen to help answer our research question is 'new housing price index'. This variable is suitable for linear regression because it is continuous, enabling us to model its relationships with multiple predictors. Additionally, since the new housing price index is influenced by economic variables, it allows us to identify trends and patterns through our linear regression model. For these reasons, it is a strong candidate for our model. Our five predictor variables are the number of available homes, homes sold, homes starting construction, new homes completed, and rate hikes

# Ethics Discussion
The datasets extracted from Statistics Canada are both trustworthy and ethically collected, as they originate from a reputable national statistical agency committed to maintaining data integrity and transparency. Statistics Canada follows strict protocols for data collection, ensuring that the information is accurate, reliable, and reflective of real market conditions (@StatCan). While the datasets themselves are typically reliable and collected following rigorous ethical standards, potential risks related to racial and ethnic disparities in housing must be acknowledged.

Ethical considerations also arise from the treatment of sensitive information, particularly regarding interest rates and housing data, which could impact individual financial circumstances. The datasets used in this study focus on aggregate data rather than individual responses, thus minimizing privacy concerns.

Moreover, the datasets underwent thorough cleaning processes to ensure accuracy, further reinforcing their reliability for analysis. Overall, the ethical collection practices and transparency associated with these datasets support their trustworthiness for examining the impacts of supply-side factors on housing prices in Toronto.

# Preliminary Results 
Focusing on the New Housing Price Index (NHPI) as the response variable, We will analyze economic, construction, and financial predictors such as construction activity, interest rates, and GDP.

## Data Preparation
```{r}
library(tidyverse)
library(here)
library(caret)
library(car)
library(corrplot)
library(broom)
library(MASS)
library(lmtest)
library(Metrics)
library(dplyr)

# Load the dataset
data <- read_csv(here::here("data/analysis_data/analysis_data.csv"))

# response variable and predictors
response <- "Quarterly_Average"
predictors <- c(
  "Detached_Absorption_Quarterly_Avg",
  "Semi_Absorption_Quarterly_Avg",
  "Detached_Unabsorbed_Quarterly_Avg",
  "Semi_Unabsorbed_Quarterly_Avg",
  "GDP_Quarterly_Avg",
  "CPI_Quarterly_Avg",
  "Starting_Detached_Construction",
  "Starting_Semi_Construction",
  "Under_Construction_Detached",
  "Under_Construction_Semi",
  "Completed_Construction_Semi",
  "Completed_Construction_Detached",
  "rates"
)

# Subset the relevant data
selected_data <- data %>%
  dplyr::select(c(response, all_of(predictors)))

# Clean Data

cleaned_data <- selected_data %>%
  drop_na()

cleaned_data <- cleaned_data %>%
  distinct()

# Initial regression model
model <- lm(as.formula(paste(response, "~", paste(predictors, collapse = " + "))), data = cleaned_data)

# Cooks Distance
cooks_d <- cooks.distance(model)

# Influential points
influential_points <- which(cooks_d > 4 / (nrow(cleaned_data) - length(model$coefficients)))

# Exclude influential observations
# cleaned_data <- cleaned_data[-influential_points, ]

# Split the Data
set.seed(123)
train_index <- createDataPartition(cleaned_data[[response]], p = 0.5, list = FALSE)
train_data <- cleaned_data[train_index, ]
test_data <- cleaned_data[-train_index, ]

list(
  training_set = train_data,
  test_set = test_data
)

```

## Assumptions and Multicollinearity

### Train Data
```{r}

# Predictors and response
response <- "Quarterly_Average"
predictors <- setdiff(names(train_data), c("Quarter", response))

# Data from the training set
eda_training_data <- train_data%>%
  dplyr::select(all_of(c(response, predictors)))

# predictors in the training set
for (predictor in predictors) {
  p <- ggplot(eda_training_data, aes_string(x = predictor, y = response)) +
    geom_point(alpha = 0.7) +
    labs(title = paste("Scatterplot of", response, "vs", predictor, "(Training Set)"),
         x = predictor, y = response) +
    theme_minimal()

  print(p)
}

# Fit the model using the training set
model_training <- lm(as.formula(paste(response, "~", paste(predictors, collapse = " + "))), data = eda_training_data)
summary(model_training)

# residuals and fitted values
residuals_training <- resid(model_training)
fitted_values_training <- fitted(model_training)

# VIF
vif_training <- vif(model_training)
print(vif_training)

# Linearity
ggplot(data = NULL, aes(x = fitted_values_training, y = residuals_training)) +
  geom_point(alpha = 0.7) +
  geom_hline(yintercept = 0, color = "red", linetype = "dashed") +
  labs(title = "Residuals vs Fitted Values",
       x = "Fitted Values",
       y = "Residuals") +
  theme_minimal()

# Residuals vs Each Predictor
for (predictor in predictors) {
  p <- ggplot(data = eda_training_data, aes_string(x = predictor, y = "residuals_training")) +
    geom_point(alpha = 0.7) +
    geom_hline(yintercept = 0, color = "red", linetype = "dashed") +
    labs(title = paste("Residuals vs", predictor),
         x = predictor, y = "Residuals") +
    theme_minimal()

  print(p)
}

# Normality of Residuals
qqnorm(residuals_training, main = "Normal Q-Q Plot")
qqline(residuals_training, col = "red")

```
### Test Data
```{r}
response <- "Quarterly_Average"
predictors <- setdiff(names(test_data), c("Quarter", response))

# test set
eda_test_data <- test_data %>%
  dplyr::select(all_of(c(response, predictors)))

for (predictor in predictors) {
  p <- ggplot(eda_test_data, aes_string(x = predictor, y = response)) +
    geom_point(alpha = 0.7) +
    labs(title = paste("Scatterplot of", response, "vs", predictor, "(Test Set)"),
         x = predictor, y = response) +
    theme_minimal()

  print(p) 
}

# Fit the model using the test set
model_test <- lm(as.formula(paste(response, "~", paste(predictors, collapse = " + "))), data = eda_test_data)
summary(model_test)

residuals_test <- resid(model_test)
fitted_values_test <- fitted(model_test)

# Calculate VIF 
vif_test <- vif(model_test)
print(vif_test)

```
The regression analysis for the **Quarterly Average** response variable demonstrates strong overall model performance. Both the training and test sets exhibit high \(R^2\) values (0.9908 and 0.9957, respectively), indicating that the models explain a significant portion of the variability in the response variable. Significant predictors include **CPI Quarterly Average**, **Detached Unabsorbed Quarterly Average**, and **Under Construction Detached**, consistently highlighting their critical impact on housing trends. However, the presence of high multicollinearity, as evidenced by elevated VIF values for variables like **CPI Quarterly Average** and **GDP Quarterly Average**, suggests a need for further refinement through dimensionality reduction or predictor selection. Residuals are well-distributed, with minimal error, confirming the robustness of the model while also indicating potential for improving simplicity and interpretability.

## Model Building

```{r}
library(car)
library(MASS)
library(lmtest)

# Transform predictors to address non-linearity
train_data$Detached_Absorption_Quarterly_Avg <- log(train_data$Detached_Absorption_Quarterly_Avg + 1)
train_data$Starting_Detached_Construction <- log(train_data$Starting_Detached_Construction + 1)
train_data$Under_Construction_Detached <- log(train_data$Under_Construction_Detached + 1)
train_data$Completed_Construction_Semi <- NULL 
train_data$Detached_Unabsorbed_Quarterly_Avg <- log(train_data$Detached_Unabsorbed_Quarterly_Avg + 1)
train_data$Semi_Unabsorbed_Quarterly_Avg <- NULL 
train_data$Starting_Semi_Construction <- NULL

# Box-Cox transformation for response variable to address skewness and heteroscedasticity
bc <- boxcox(Quarterly_Average ~ Detached_Absorption_Quarterly_Avg +
                                Starting_Detached_Construction +
                                Detached_Unabsorbed_Quarterly_Avg +
                                Under_Construction_Detached +
                                CPI_Quarterly_Avg,
             data = train_data, plotit = FALSE)
lambda <- bc$x[which.max(bc$y)]

if (lambda == 0) {
  train_data$Quarterly_Average <- log(train_data$Quarterly_Average)
} else {
  train_data$Quarterly_Average <- (train_data$Quarterly_Average^lambda - 1) / lambda
}

# Fit the linear regression model
model <- lm(Quarterly_Average ~ Detached_Absorption_Quarterly_Avg +
                                   Starting_Detached_Construction +
                                   Detached_Unabsorbed_Quarterly_Avg +
                                   Under_Construction_Detached +
                                   CPI_Quarterly_Avg,
            data = train_data)

```
The transformations applied in the code aim to address key violations of regression assumptions identified in the data. Logarithmic transformations were used for predictors such as `Detached_Absorption_Quarterly_Avg`, `Starting_Detached_Construction`, and `Detached_Unabsorbed_Quarterly_Avg` to address non-linearity observed in the scatterplots, ensuring a more linear relationship with the response variable, `Quarterly_Average`. Predictors like `Starting_Semi_Construction` and `Semi_Unabsorbed_Quarterly_Avg` were removed due to their highly non-linear behavior, which made them less suitable for inclusion in the model. Similarly, the Box-Cox transformation was applied to the response variable to correct skewness and heteroscedasticity, ensuring constant variance and improving model accuracy. These adjustments collectively enhance the model's adherence to assumptions, reduce potential bias, and improve interpretability.

```{r}

predictors <- c(
  "Detached_Absorption_Quarterly_Avg",
  "Detached_Unabsorbed_Quarterly_Avg",
  "GDP_Quarterly_Avg",
  "CPI_Quarterly_Avg",
  "Starting_Detached_Construction",
  "Under_Construction_Detached",
  "Completed_Construction_Detached",
  "rates"
)

# Residuals and fitted values
residuals_model <- resid(model)
fitted_values_model <- fitted(model)

# VIF
vif_model <- vif(model)
print(vif_model)

# Linearity
ggplot(data = NULL, aes(x = fitted_values_model, y = residuals_model)) +
  geom_point(alpha = 0.7) +
  geom_hline(yintercept = 0, color = "red", linetype = "dashed") +
  labs(title = "Residuals vs Fitted Values",
       x = "Fitted Values",
       y = "Residuals") +
  theme_minimal()

# Residuals vs Each Predictor
for (predictor in predictors) {
  p <- ggplot(data = train_data, aes_string(x = predictor, y = "residuals_model")) +
    geom_point(alpha = 0.7) +
    geom_hline(yintercept = 0, color = "red", linetype = "dashed") +
    labs(title = paste("Residuals vs", predictor),
         x = predictor, y = "Residuals") +
    theme_minimal()

  print(p)
}

# Normality of Residuals
qqnorm(residuals_model, main = "Normal Q-Q Plot")
qqline(residuals_model, col = "red")

```

### Assumptions and Multicollinearity
Overall, the model's assumptions are reasonably satisfied. Minor issues with linearity and normality at the tails could be further investigated, but they are not critical. The transformations applied to some predictors appear effective in addressing non-linearity and heteroscedasticity.

## Model Diagnostics
```{r}
# Model Performance Metrics
model_summary <- summary(model)

# Extract R^2 and Adjusted R^2
r_squared <- model_summary$r.squared
adj_r_squared <- model_summary$adj.r.squared

# AIC and BIC
aic <- AIC(model)
bic <- BIC(model)

# RMSE
residuals <- residuals(model)
rmse <- sqrt(mean(residuals^2))

# Results
cat("Model Performance Metrics:\n")
cat("R-squared: ", round(r_squared, 4), "\n")
cat("Adjusted R-squared: ", round(adj_r_squared, 4), "\n")
cat("AIC: ", round(aic, 4), "\n")
cat("BIC: ", round(bic, 4), "\n")
cat("RMSE: ", round(rmse, 4), "\n")

```

The regression model demonstrates a strong overall fit, with an R-squared of 0.9832 and an Adjusted R-squared of 0.9807, indicating that the predictors explain a significant portion of the response variable's variability while maintaining robustness. The low RMSE (5.4798) reflects good predictive accuracy, and the AIC (263.6004) and BIC (275.4225) suggest a reasonable balance between model complexity and fit. These metrics highlight the model's effectiveness, though addressing minor residual patterns could further enhance its performance.

## 6. Model Validation

```{r}

# Predict on test_data
predictions <- predict(model, newdata = test_data)

# Calculate residuals
residuals <- test_data$Quarterly_Average - predictions

# Compute MSE for test_data
mse_test <- mean(residuals^2)
cat("Mean Squared Error (MSE) on Test Data:", mse_test, "\n")

# R-squared for test_data
ss_total <- sum((test_data$Quarterly_Average - mean(test_data$Quarterly_Average))^2)
ss_residual <- sum(residuals^2)
r_squared_test <- 1 - (ss_residual / ss_total)
cat("R-squared on Test Data:", r_squared_test, "\n")

# Print validation results
cat("Validation Results:\n")
cat("Mean Squared Error (Test Data):", mse_test, "\n")
cat("R-squared (Test Data):", r_squared_test, "\n")

```
The model's performance on the test data indicates significant issues. The **Mean Squared Error (MSE)** is exceptionally high (\(1.303812 \times 10^{15}\)), demonstrating substantial deviations between predicted and actual values. Additionally, the **R-squared** is highly negative (\(-4.163317 \times 10^{12}\)), indicating that the model performs far worse than simply using the mean of the response variable. These results suggest the model fails to generalize effectively to unseen data.

```{r}
saveRDS(model, file = here("data", "final_model.rds"))
```

\newpage

# Bibliography